Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra using Murchison Widefield Array Observations

### 2015-04-07 ms.tex
Modify Abstract
From e637668:
============

In ms.tex,

(*) Modify abstract

This analysis confirms recent predictions of certain key signatures in the delay spectra of wide-field radio interferometer measurements. While probing H~{\sc i} in the epoch of reionization (EoR) through modeling of delay spectra of measurements between antenna pairs, it has recently emerged that the foreground structure, the nature of wide-field measurements and the antenna aperture imprint a characteristic {\it pitchfork}-shaped signature. This feature is characterized by enhanced power from foreground emission in Fourier modes mapped to regions near the horizon, and is significant for EoR studies because Fourier modes considered to be most sensitive for the cosmological H~{\sc i} signal are very susceptible to contamination from these regions. Thus it forms a critical input to design and analysis choices of future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array. By improving Murchison Widefield Array data sensitivity through coherent averaging of 12 independent snapshots aligned in local sidereal time across different observing nights, we provide the first confirmation of the prediction with a signal-noise ratio $>10$.

@@

The epoch of reionization (EoR) commenced following the formation of the first stars and galaxies. It is characterized by a period of non-linear growth of matter density perturbations and astrophysical evolution in the Universe's history. Detection of redshifted 21~cm radiation of H~{\sc i} from this epoch is one of the most promising probes of the non-linear evolution of large scale structure characteristic of this epoch \citep{sun72,sco90,mad97,toz00,ili02}.

>>>>

Here we confirm a recently predicted and previously unknown foreground signature in the 3D power spectra of high redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherently chromatic nature of a wide-field instrument response and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. Thus it is a critical input to design and analysis choices of future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array. The simulation which predicted this feature was recently validated against Murchison Widefield Array data but this key element was at or below the noise level. In this paper, we improve the Murchison Widefield Array data sensitivity through coherent averaging of 12 independent snapshots aligned in local sidereal time across different observing nights, and provide the first confirmation of the prediction with a signal-noise ratio $>10$.

% This analysis confirms recent predictions of certain key signatures in the delay spectra of wide-field radio interferometer measurements. While probing H~{\sc i} in the epoch of reionization (EoR) through modeling of delay spectra of measurements between antenna pairs, it has recently emerged that the foreground structure, the nature of wide-field measurements and the antenna aperture imprint a characteristic {\it pitchfork}-shaped signature. This feature is characterized by enhanced power from foreground emission in Fourier modes mapped to regions near the horizon, and is significant for EoR studies because Fourier modes considered to be most sensitive for the cosmological H~{\sc i} signal are very susceptible to contamination from these regions. Thus it forms a critical input to design and analysis choices of future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array. By improving Murchison Widefield Array data sensitivity through coherent averaging of 12 independent snapshots aligned in local sidereal time across different observing nights, we provide the first confirmation of the prediction with a signal-noise ratio $>10$.

@@

The epoch of reionization (EoR) commenced following the formation of the first stars and galaxies. It is characterized by a period of non-linear growth of matter density perturbations and astrophysical evolution in the Universe's history. Detection of redshifted 21~cm radiation of H~{\sc i} from this epoch is one of the most promising probes of the evolution of large scale structure characteristic of this epoch \citep{sun72,sco90,mad97,toz00,ili02}.
